Toronto Island Ferry Tickets: Sales, Redemption Analysis, and Strategic Solutions

### 2024-09-27 paper/paper.qmd
update README, LLM usage and added documentation to scripts

---
title: "Patterns of Toronto Islands Ferries Tickets Sales and Redemption and potential Solutions"
subtitle: ""
author: 
  - Yun Chu
thanks: "code and data supporting this analaysis can be found here: https://github.com/chuyun2024/Analyzing-Toronto-Islands-Ferry-Tickets-Sales-and-Redemption-Count-Data"
date: 26 September 2024
date-format: long
abstract: "The paper firstly discussed the background of the dataset, then presented and analyzed the cleaned dataset. In the result section, the relationship between the sales and redemption data, the distribution of redemption count, and the trend of redemption count with respect to time is analyzed. The discussion section comes up with some suggestions to Toronto Islands ferries based on the findings and discuss the limilation of the analysis."
format: pdf
toc: true
number-sections: true
bibliography: references.bib
---

@@

# Create a formatted table to display the sample data
# kable generates the table, with specified column names and alignment
# 'align = "ccc"' aligns all three columns to the center
# caption adds a descriptive title to the table
kable(data2, 
      col.names = c('Date', 'Sales', 'Redemption'), 
      align = "ccc",  
      caption = "Sample of Cleaned Sales and Redemption Data")
```

@@

# Create a line plot using ggplot2
ggplot(data, aes(x = formatted_date)) +  # Define the data and aesthetic mappings; formatted_date on the x-axis
  geom_line(aes(y = daily_sales_count, color = "Sales"), linewidth = 1, alpha = 0.7) +  # Add a line for daily sales count
  geom_line(aes(y = daily_redemption_count, color = "Redemption"), linewidth = 1, alpha = 0.7) +  # Add a line for daily redemption count
  labs(title = "Sales and Redemption Over Time",  # Add titles and labels to the plot
       x = "Date",                                # Label for the x-axis
       y = "Count") +                             # Label for the y-axis
  scale_color_manual(name = "Legend",             # Customize the color scale for the legend
                     values = c("Sales" = "#2b5b84", "Redemption" = "#98df8a")) +  # Assign specific colors to the lines
  theme_minimal() +                               # Use a minimal theme for cleaner visuals
  theme(plot.title = element_text(hjust = 0.5))  # Center the plot title horizontally
```

@@

```{r}
#| label: tbl-mean_sales_and_redemption
#| echo: false
#| warning: false
#| message: false
# Calculate summary statistics
summary_statistic <- data |>
  summarise(
    Min_Sales = min(daily_sales_count),
    Max_Sales = max(daily_sales_count),
    Mean_Sales = mean(daily_sales_count),
    Std_Sales = sd(daily_sales_count),
    Min_Redemption = min(daily_redemption_count),
    Max_Redemption = max(daily_redemption_count),
    Mean_Redemption = mean(daily_redemption_count),
    Std_Redemption = sd(daily_redemption_count))

@@

# Display the summary table with row and column headings
kable(summary_table, format = "latex", row.names = TRUE, col.names = c("Min", "Max", "Mean", "Standard Deviation"), 
      caption = 'Summary Statistics for Sales and Redemption Count',
      align = "ccccc") |>
  kable_styling(full_width = FALSE, position = "left")
```

@@

# Create a scatter plot using ggplot2 to visualize the relationship between sales and redemption
ggplot(data, aes(x = daily_sales_count, y = daily_redemption_count)) +  # Set the data and aesthetic mappings; daily sales on the x-axis and daily redemption on the y-axis
  geom_point(alpha = 0.5, color = "#008080") +  # Add points to the plot to represent each observation
  labs(title = "Scatter Plot of Sales vs Redemption",  # Add a title to the plot
       x = "Ticket Sales Count",  # Label for the x-axis
       y = "Ticket Redemption Count") + # Label for the y-axis
  theme(plot.title = element_text(hjust = 0.5))
```

@@

# Render the summary table using kable for better formatting, with specified column names and caption
kable(summary_table, col.names = c("Statistics", "Value"), align = 'cc', caption = "R-squared and Correlation Statistic")
```

@@

# Create a histogram to visualize the distribution of daily redemption counts
ggplot(data, aes(x = daily_redemption_count)) +  
  geom_histogram(binwidth = 300, fill = "lightskyblue", color = "steelblue1") +  # Add histogram bars with specified binwidth and colors
  labs(title = "Histogram of Daily Redemption Counts",  # Add a title for the histogram
       x = "Daily Redemption Count",  # Label for the x-axis
       y = "Frequency") +  # Label for the y-axis
  theme(plot.title = element_text(hjust = 0.5)) +  # Center the plot title
  annotate("text", x = percentile_95, y = 30, label = paste("95th Percentile:", round(percentile_95, 2)),  # Annotate the plot with the 95th percentile value
           color = "mediumpurple", vjust = -0.5) +  # Set the color of the annotation and adjust vertical positioning
  geom_vline(xintercept = percentile_95, color = "mediumpurple", linetype = "dashed", size = 1)  # Add a dashed vertical line at the 95th percentile

```

@@

# Create a bar plot to visualize redemption counts over each month
ggplot(data, aes(x = year_month, y = daily_redemption_count)) +  
  geom_bar(stat = "identity", fill = "grey93", color = "lightskyblue") +  # Create bars with specified fill and border colors
  labs(title = "Redemption Over Month",  # Add a title for the plot
       x = "Month",  # Label for the x-axis
       y = "Redemption Count") +  # Label for the y-axis
  theme(axis.text.x = element_text(angle = 45, hjust = 1),  # Rotate x-axis text for better readability
        plot.title = element_text(hjust = 0.5)) +  # Center the plot title
  scale_y_continuous(labels = comma)  # Format the y-axis labels to include commas for thousands

```

@@

Based on @fig-Daily_Redemption_Histogram, one measure that Toronto Islands ferries could implement is to integrate staff schedules during peak seasons to prepare for large visitor volumes. For instance, having janitors on-call during busy periods would ensure that adequate cleaning services are available when the number of visitors increases significantly.

Additionally, it is essential to develop an on-call schedule that aligns with peak and off-peak seasons. For example, in the winter months, preparing a full staff contingent to manage more than 20,000 visitors per day would be an inefficient use of resources. Instead, a more tailored approach would allow for better resource allocation based on anticipated demand.

## Evaluation of Staffing Schedules in Relation to Visitor Demand

@@

Given the extremely high volume of visitors in July and August, adjusting the staffing schedules during these peak months may be more beneficial. This approach would align more closely with visitor demand and help avoid resource wastage during periods of lower activity.

>>>>

---
title: "Toronto Island Ferry Tickets: Sales, Redemption Analysis, and Strategic Solutions"
subtitle: ""
author: 
  - Yun Chu
thanks: "code and data supporting this analaysis can be found here: https://github.com/chuyun2024/Analyzing-Toronto-Islands-Ferry-Tickets-Sales-and-Redemption-Count-Data"
date: 26 September 2024
date-format: long
abstract: "This paper presents a comprehensive analysis of the Toronto Islands ferry ticket sales and redemption dataset. The analysis begins with an overview of the dataset, followed by an examination of the cleaned data. In the results section, we explore the relationship between sales and redemption data, the distribution of redemption counts, and the monthly trends in redemption activity. Based on the findings, we offer recommendations to optimize ferry operations, focusing on pricing strategies, staffing allocations, and departure schedules. The paper concludes by discussing the limitations of the analysis and identifying potential areas for future improvement."
format: pdf
toc: true
number-sections: true
bibliography: references.bib
---

@@

# Create a formatted table to display the sample data
# kable generates the table, with specified column names and alignment
# 'align = "ccc"' aligns all three columns to the center
# caption adds a descriptive title to the table
kable(data2,
  col.names = c("Date", "Sales", "Redemption"),
  align = "ccc",
  caption = "Sample of Cleaned Sales and Redemption Data"
)
```

@@

# Create a line plot using ggplot2
ggplot(data, aes(x = formatted_date)) + # Define the data and aesthetic mappings; formatted_date on the x-axis
  geom_line(aes(y = daily_sales_count, color = "Sales"), linewidth = 1, alpha = 0.7) + # Add a line for daily sales count
  geom_line(aes(y = daily_redemption_count, color = "Redemption"), linewidth = 1, alpha = 0.7) + # Add a line for daily redemption count
  labs(
    title = "Sales and Redemption Over Time", # Add titles and labels to the plot
    x = "Date", # Label for the x-axis
    y = "Count"
  ) + # Label for the y-axis
  scale_color_manual(
    name = "Legend", # Customize the color scale for the legend
    values = c("Sales" = "#2b5b84", "Redemption" = "#98df8a")
  ) + # Assign specific colors to the lines
  theme_minimal() + # Use a minimal theme for cleaner visuals
  theme(plot.title = element_text(hjust = 0.5)) # Center the plot title horizontally
```

@@

```{r}
#| label: tbl-mean_sales_and_redemption
#| echo: false
#| warning: false
#| message: false
# Calculate summary statistics
summary_statistic <- data |>
  summarise(
    Min_Sales = min(daily_sales_count),
    Max_Sales = max(daily_sales_count),
    Mean_Sales = mean(daily_sales_count),
    Std_Sales = sd(daily_sales_count),
    Min_Redemption = min(daily_redemption_count),
    Max_Redemption = max(daily_redemption_count),
    Mean_Redemption = mean(daily_redemption_count),
    Std_Redemption = sd(daily_redemption_count)
  )

@@

# Display the summary table with row and column headings
kable(summary_table,
  format = "latex", row.names = TRUE, col.names = c("Min", "Max", "Mean", "Standard Deviation"),
  caption = "Summary Statistics for Sales and Redemption Count",
  align = "ccccc"
) |>
  kable_styling(full_width = FALSE, position = "left")
```

@@

# Create a scatter plot using ggplot2 to visualize the relationship between sales and redemption
ggplot(data, aes(x = daily_sales_count, y = daily_redemption_count)) + # Set the data and aesthetic mappings; daily sales on the x-axis and daily redemption on the y-axis
  geom_point(alpha = 0.5, color = "#008080") + # Add points to the plot to represent each observation
  labs(
    title = "Scatter Plot of Sales vs Redemption", # Add a title to the plot
    x = "Ticket Sales Count", # Label for the x-axis
    y = "Ticket Redemption Count"
  ) + # Label for the y-axis
  theme(plot.title = element_text(hjust = 0.5))
```

@@

# Render the summary table using kable for better formatting, with specified column names and caption
kable(summary_table, col.names = c("Statistics", "Value"), align = "cc", caption = "R-squared and Correlation Statistic")
```

@@

# Create a histogram to visualize the distribution of daily redemption counts
ggplot(data, aes(x = daily_redemption_count)) +
  geom_histogram(binwidth = 300, fill = "lightskyblue", color = "steelblue1") + # Add histogram bars with specified binwidth and colors
  labs(
    title = "Histogram of Daily Redemption Counts", # Add a title for the histogram
    x = "Daily Redemption Count", # Label for the x-axis
    y = "Frequency"
  ) + # Label for the y-axis
  theme(plot.title = element_text(hjust = 0.5)) + # Center the plot title
  annotate("text",
    x = percentile_95, y = 30, label = paste("95th Percentile:", round(percentile_95, 2)), # Annotate the plot with the 95th percentile value
    color = "mediumpurple", vjust = -0.5
  ) + # Set the color of the annotation and adjust vertical positioning
  geom_vline(xintercept = percentile_95, color = "mediumpurple", linetype = "dashed", size = 1) # Add a dashed vertical line at the 95th percentile
```

@@

# Create a bar plot to visualize redemption counts over each month
ggplot(data, aes(x = year_month, y = daily_redemption_count)) +
  geom_bar(stat = "identity", fill = "grey93", color = "lightskyblue") + # Create bars with specified fill and border colors
  labs(
    title = "Redemption Over Month", # Add a title for the plot
    x = "Month", # Label for the x-axis
    y = "Redemption Count"
  ) + # Label for the y-axis
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1), # Rotate x-axis text for better readability
    plot.title = element_text(hjust = 0.5)
  ) + # Center the plot title
  scale_y_continuous(labels = comma) # Format the y-axis labels to include commas for thousands
```

@@

Based on @fig-Daily_Redemption_Histogram, one measure that Toronto Islands ferries could implement is to integrate staff and departure schedules during peak seasons to prepare for large visitor volumes. For instance, having janitors on-call during busy periods would ensure that adequate cleaning services are available and shortening the time interval between departures when the number of visitors increases significantly.

Additionally, it is essential to develop an on-call schedule for staff that aligns with peak and off-peak seasons. For example, in the winter months, preparing a full staff contingent to manage more than 20,000 visitors per day would be an inefficient use of resources. Instead, a more tailored approach would allow for better resource allocation based on anticipated demand.

## Evaluation of Staffing and Departure Schedules in Relation to Visitor Demand

@@

Given the extremely high volume of visitors in July and August, adjusting the staffing and departure schedules during these peak months may be more beneficial. This approach would align more closely with visitor demand and help avoid resource wastage during periods of lower activity.
